---
title: "math 545 asst 4"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(itsmr)
library(forecast)
library(tidyverse)
```
## 6.9

### a.)

#First, we will difference the data. Based on the yearly structure of the data
#we will difference with a lag of 12 in our model. Then we will check for our orders o


```{r cars}
beer_dat = dget("beer.Rput")
#log the data and remove the last 12 points
beer_short<-ts(beer_dat[1:(length(beer_dat)-12)], frequency = 12)
beer_short<-log(beer_short)

#check the arima models
data_arima = auto.arima(beer_short, seasonal = TRUE, stepwise=FALSE, approximation = FALSE)
summary(data_arima)

beer_arima1 <- auto.arima(beer_short, seasonal=TRUE, stepwise=TRUE,
                          approximation=FALSE, allowdrift=FALSE)

summary(beer_arima1)
## This model performs well, as an ARIMA(0,1,3)(0,1,2)[12] seasonal arima model

## However I have found a model which has a more better AIC.
beer_arima2 <- arima(beer_short, c(13, 1, 6))
summary(beer_arima2)
```

##b

```{r 6.9b}
## Confidence interval for beer_arima1 model
confint(beer_arima1)
```

### c.)

```{r 6.9c}
resid <- beer_arima1$residuals

## residuals
plot(resid)

## Distribution of residuals
hist(resid)

## Acf of residuals
ggAcf(resid)
```

the residual patterns look to be approximately normal and approximately constant variance, and centered around zero. It is not clear if they are stationary.

### d.) and e.)

```{r 6.9de}
frcst <- forecast(beer_arima1, h=12, level=95)
## Forecasted values with the first model
plot(frcst)

## Forecasts and 95% prediction bounds
f1 <- exp(as.data.frame(frcst))
names(frcst) <- gsub(" ", "", paste0("model1", names(frcst)))
f1
```
``` {r 6.9f, results = FALSE, include = FALSE}
fcast2 <- forecast(beer_arima2, h=12, level=95)
## Forecasted values with the first model
plot(fcast2)
## Forecasts and 95% prediction bounds for second model
f2 <- exp(as.data.frame(fcast2))
names(f2) <- gsub(" ", "", paste0("model2", names(f2)))
f2
```

### f.) 

``` {r 6.9f}
df <- data.frame("beer"=beer_dat[(length(beer_dat)-11):length(beer_dat)]) %>% 
  cbind(f1, f2) %>%
  mutate(model1error=beer-model1PointForecast, model2error=beer-model2PointForecast) %>%
  select(model1error, model2error)
df
```

## 6.10
```{r 6.10a}
beer <- dget("beer.Rput")


## Log of data, minus the last 12 samples
beer_short <- data.frame('time'=1:(length(beer)-12), 'beer'=log(ts(beer[1:(length(beer) - 12)], frequency=12)))

beer_short <- beer_short %>% mutate(seasonal = season(beer, d=12)) %>%
  mutate(seasonal_resid = beer - seasonal,
         quad_time = time^2)
quad_trend <- lm(data=beer_short, seasonal_resid~time + quad_time)
beer_short <- beer_short %>% 
  mutate(seas_quad = quad_trend$fitted.values) %>%
  mutate(seas_quad_resid = seasonal_resid - seas_quad,
         preds = seas_quad + seasonal)

ggAcf(beer_short$seas_quad_resid)

## Same method as previous question
beer_arima <- auto.arima(beer_short$seas_quad_resid, 0, 0, max.order = 20, seasonal=FALSE, stepwise=FALSE,
                         approximation=FALSE, allowdrift=FALSE)


```


### b.)

```{r 6.10b}
## Confidence interval
confint(beer_arima)
```

### c.)

```{r 6.10c}
beer_resid <- beer_arima$residuals
## Residuals
plot(beer_resid)

## Distribution of model residuals
hist(beer_resid)

## Acf of model residuals
ggAcf(beer_resid)

```
The residual look normally distributed, centered around zero, and with constant variance, but they may not satisfy the stationary assumption

### d.) and e.)

```{r 6.10de}
beer_short_new <- data.frame('seasonal' = beer_short$seasonal_resid[(nrow(beer_short) - 11):nrow(beer_short)],
                             'time'=411:422) %>% 
  mutate('quad_time' = time^2) 
beer_short_new <- beer_short_new %>%
  mutate('quad_preds' = predict(quad_trend, beer_short_new))

fcast1 <- forecast(beer_arima, h=12, level=95)
f1 <- exp(as.data.frame(fcast1) + beer_short_new$quad_preds)
names(f1) <- gsub(" ", "", names(f1))


## Forecasted residuals from seasonal + quad
plot(fcast1)

length((((1991 + 3/12) * 12):((1992 + 2/12) * 12))/12)
## Forecasted actual values
autoplot(exp(beer_short$beer)) + geom_line(data=f1, aes(x=((length(beer)):(length(beer)+11))/12, y=PointForecast), colour='blue') + 
  geom_ribbon(data=f1, aes(x=((length(beer)):(length(beer)+11))/12, ymin=Lo95, ymax=Hi95), alpha=0.3, inherit.aes = FALSE)
     
              
f1
```

### f.) 

``` {r 6.10f}

df <- data.frame("beer"=beer[(length(beer)-11):length(beer)]) %>% 
  cbind(f1) %>%
  mutate(error=beer-PointForecast)
df
```


The last value (as shown in part e) is within the boundaries. These errors are much larger, and the model tends to do a much poorer job of catching the residuals from the original dataset than the results from last question. 